---
title: "Customer Churn Analysis"
author: "YOUR NAME HERE"
date: "7/20/2021"
output: html_document
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## libraries
```

## Data Read

```{r}
churn <- read_csv(here("Data", "________"))
```

```{r}
str(churn)
summary(churn)
glimpse(churn)
```


## Data Exploration

```{r}
churn %>%
  ggplot() +
  geom_histogram(aes(x = Tenure))

## Explore more quantitative variables
```


```{r}
## Explore some categorical or discrete variables

churn %>%
  count(_____)

churn %>%
  ggplot() +
  geom_bar(aes(x = _______))

```

## Data Cleaning


```{r}
churn <- churn %>%
  mutate(
    ### make necessary variables into factors
  )
```

## Data Summaries

1. What are the mean and median account balances among customers in this dataset?

2. What are the mean values of: age, account balance, and salary?

3. Among customers in the USA, what is the average credit score?

4. What country are the customers with the top three highest balances from?

5. Do the average salaries differe between those with credit cards and those without?

6. Is there a difference in gender distribution of customers who have exited the bank?


## More Visualization

```{r}
## Comparing categorical variables

churn %>%
  ggplot() +
  geom_bar(aes(x = ______,
               fill = ______))

churn %>%
  ggplot() +
  geom_bar(aes(x = ______,
               fill = ______),
           position = "fill")


churn %>%
  ggplot() +
  geom_bar(aes(x = ______,
               fill = ______),
           position = "dodge")


churn %>%
  ggplot() +
  geom_bar(aes(x = ______,
               fill = ______)) +
  facet_wrap(~_______)
```

```{r}
## comparing a categorical and a quantitative

________ +
  geom_boxplot(______)


_______ + 
  geom_density(________,
               alpha = 0.5)

```


```{r}
## comparing two quantitatives

#### Find the geometry for a scatterplot

```


## Pivoting

1. Find the average salary of each gender in each country.

2. Pivot your data frame from (1) so that there is one column for each gender identity.

3. Create a new column showing the difference between female-identifying and male-identifying average salaries.


## Joining

Repeat the above, this time for balances rather than salaries.

Join the two datasets by country name.


## Your own analysis

Choose one of the following research questions to address with a beautiful plot (or make up your own).  Write up your results in a **new** R Markdown file, in your **new** project.

* How does the relationship between age and churn (exit) differ by gender/country?

* Is a customer with a credit card considered "desirable"?  For example, do they stay with the bank longer?  Are they less likely to exit? Do they use more bank products?

* How does age impact customer behavior?  That is, is the impact of other qualities (balance, salary, etc) on customer tenure and/or churn different across ages?

