---
title: "Packages and Projects"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
solarized_light()
```

## Principles of Reproducibility

The idea:  You should be able to send your project to someone else, and they can jump in and start working right away.

1. Files are organized and well-named.

2. References to data and code work for everyone.

3. Package dependency is clear.

4. Code will run the same every time, even if data values change.

5. Analysis process is well-explained and easy to read.

```{r, include = FALSE}
library(tidyverse)
```

---
# Dealing with Directories


- Your **working directory** is the folder that R "thinks" it lives in at the moment.

```{r getwd}
getwd()
```

- If you save things you have created, they save to your working directory by default.


---

## The beauty of R Projects

An **R Project** is basically a "flag" planted in a certain directory.

When you double click a **.Rproj** file, it does two things:

1. Open RStudio

2. Set the working directory to be wherever the .Rproj file lives.

3. Link to GitHub, if setup.

--
This is great for reproducibility!  

Anyone can copy your directory, and the .Rproj will open properly on their computer.

---
class: center, middle

# Set up your RProj

In RStudio, choose `File > New Project`.

(Save any files you need to.)

Choose **new directory** and create a new folder.

--

Close **all** of your RStudio windows.

Go to your new folder.

Double click the **.Rproj** file (it looks like a cube!)

--

**You should ALWAYS access your work this way from now on!**

---
# Bad practice

If you put something like this at the top of your .Rmd file, that is not great:

```{r setwd, eval = FALSE}
setwd("/User/kbodwin/Dropbox/Teaching/MACU_Workshop/Data/")
```

- Setting working directory by hand = BAD!  

--

- That directory is specific to you!  Where should someone else look for the data?

--

- Markdown files won't knit this code.

---
class: center, middle, inverse

# Dealing with packages

---

## install.packages()

To use a package, it must first be *installed* to your copy of RStudio.

This only needs to happen **ONCE** per computer or server.

```{r install packages, eval = FALSE}
install.packages("tidyverse")
```

--

## update.packages()

If a computer you are using has the package, but you aren't sure if it's up to date:

```{r update packages, eval = FALSE}
update.packages("tidyverse")
```

to install the most recent version.

---
class: center,middle
# NEVER EVER EVER EVER PUT "install.packages" or "update.packages" INTO YOUR R MARKDOWN FILE!!!!!

---

# library() and require()

- A package must only be *installed* once.

- But it must be *loaded* every time, to tell R it is needed for this analysis.

- `library` gives an **error** if the package isn't installed

- `require` gives a `TRUE/FALSE` value depending on if the package was available

- For now, you should always use `library`

```{r library, eval = FALSE}
library(tidyverse)

# notice:  no quotes around package name
```

---
class: center, middle

# Three annoying complications of packages

---

## Versions

- From time to time, packages get updated

- This is good!

- But, it can mess with your code, which depended on the older version.

- No magic answer; just be aware.

- Typically, 

    + package 1.0.1 to 1.0.2 is a non-breaking change
    + package 1.1.0 to 1.2.0 is a non-breaking change with warnings
    + package 1.0.0 to 2.0.0 is a breaking change


---

## Dependencies

- A *dependency* is when one package makes use of another.

- For example, if you create a package that uses the function `group_by()`, your package **depends** on `dplyr`

- Typically these get loaded automatically!

- Occasionally, something goes weird and you need to manually install the dependency.


---


## Conflicts

- Sometimes, packages choose the same name for their functions.  This is called a *conflict*.

- The version loaded first takes precedence.  For example...

```{r conflicts1, message = FALSE}
dat <- data.frame(bob = 1:10, joe = 11:20)

library(MASS)
library(dplyr)
```

```{r, error = TRUE}
dat %>% select(bob)
```

---
## Conflicts

```{r, echo = FALSE, message = FALSE, warning = FALSE}
detach("package:dplyr", unload=TRUE)
detach("package:MASS", unload=TRUE)
```

```{r conflicts2, message = FALSE}
dat <- data.frame(bob = 1:10, joe = 11:20)

library(dplyr)
library(MASS)
```

```{r, error = TRUE}
dat %>% select(bob)
```

---

## Conflicts

But this is super annoying, because...

```{r, echo = FALSE, message = FALSE, warning = FALSE}
detach("package:dplyr", unload=TRUE)
detach("package:MASS", unload=TRUE)
```

```{r conflicts3, eval = FALSE}
dat <- data.frame(bob = 1:10, joe = 11:20)

library(dplyr)  # dplyr loaded here
library(MASS)
library(dplyr)  # does nothing, because it sees we already have dplyr
```

```{r, error = TRUE}
dat %>% select(bob)
```

---

# Conflicts

ARG!

Best solution:  Be explicit about your intended package using `::`.

```{r conflicts4, message = FALSE, eval = FALSE}
dat <- data.frame(bob = 1:10, joe = 11:20)

library(dplyr) 
library(MASS)

dat %>% dplyr::select(bob)
```

--

Another solution: The `conflicted` package

```{r}
library(conflicted)

conflict_prefer("select", "dplyr")
```

---
class: center, middle, invert
# Deliverables in R Markdown

---
# General principles

- Give your file a good title, author, and date

- Name your file something meaningful (match the title!)

- Use `#` to make section headings etc.

- Each code chunk is a unified thought.

- Text before and after each code chunk to motivate your code and explain the output.

- Make deliberate decisions about what to display:
    + Does the reader need to see data loading step?
    + Does the reader need to see your code?
    + Does the reader need to see the output?

---

---  

## Naming code chunks

- Lets you jump between chunks

- Careful with copy-pasting, duplicate names are not allowed.

(demo)

---
# Setting chunk options

- `echo`: Should this code be **shown** (echoed)?

- `eval`: Should this code be **run** (evaluated)?

- `include`: Should this code be neither shown nor run?

- `error`: If there is an error, should that display rather than breaking the knit?

- `warning`: If there is a warning, should that display?

- `message`: If there is a message, should that display?

---
## Setting options for ALL chunks

What does this do?

```{r global_options, eval=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4,
                      echo=TRUE, warning=FALSE, 
                      message=FALSE, eval=TRUE)
```


- (you can always override these for individual chunks)
